---
title: "Predict GEBV for NRCRI C2"
author: "wolfemd"
date: "2020-April-27"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = F, tidy = T)
```

# Previous step

4. [Check prediction accuracy](04-CrossValidation.html): Evaluate prediction accuracy with cross-validation.

# Objective

**Current Step**  

5. [Genomic prediction of GS C2](05-GetGBLUPs.html): Predict _genomic_ BLUPs (GEBV and GETGV) for all selection candidates using all available data.

```{bash, eval=F}
# activate multithread OpenBLAS 
export OMP_NUM_THREADS=48
#export OMP_NUM_THREADS=88
#export OMP_NUM_THREADS=88
```

# Set-up training-testing data

```{r, eval=F}
library(tidyverse); library(magrittr); 
A<-readRDS(file=here::here("output","Kinship_A_NRCRI_2020April27.rds"))
D<-readRDS(file=here::here("output","Kinship_A_NRCRI_2020April27.rds"))
#AA<-readRDS(file=here::here("output","Kinship_AA_NRCRI_2020April27.rds"))
AD<-readRDS(file=here::here("output","Kinship_AD_NRCRI_2020April27.rds"))
#DD<-readRDS(file=here::here("output","Kinship_DD_NRCRI_2020April27.rds"))

blups_iita<-readRDS(file=here::here("data","iita_blupsForCrossVal_outliersRemoved_73019.rds"))
blups_iita<-blups_iita %>% 
  dplyr::select(Trait,blups) %>% 
  unnest(blups) %>% 
  dplyr::select(-`std error`) %>% 
  filter(GID %in% rownames(A),
         !grepl("TMS13F|TMS14F|TMS15F|2013_",GID)) # don't want IITA GS progenies
blups_nrcri<-readRDS(file=here::here("output","nrcri_blupsForModelTraining_2020April27.rds"))
blups_nrcri<-blups_nrcri %>% 
  dplyr::select(Trait,modelOutput) %>% 
  unnest(modelOutput) %>% 
  dplyr::select(Trait,BLUPs) %>% 
  unnest(BLUPs) %>%  
  filter(GID %in% rownames(A))

blups<-blups_nrcri %>%
  nest(TrainingData=-Trait) %>% 
  mutate(Dataset="NRCRIalone") %>% 
  bind_rows(blups_nrcri %>% 
              bind_rows(blups_iita %>% filter(Trait %in% blups_nrcri$Trait)) %>% 
              nest(TrainingData=-Trait) %>% 
              mutate(Dataset="IITAaugmented"))
blups
```

# Prediction

## runGenomicPredictions function
```{r runGenomicPredictions, eval=F}
#' @param blups nested data.frame with list-column "TrainingData" containing BLUPs
#' @param modelType string, A, AD or ADE representing model with Additive-only, Add. plus Dominance, and Add. plus Dom. plus. Epistasis (AA+AD+DD), respectively.
#' @param grms list of GRMs. Any genotypes in the GRMs get predicted with, or without phenotypes. Each element is named either A, D, AA, AD, DD. Matrices supplied must match required by A, AD and ADE models. For ADE grms=list(A=A,D=D,AA=AA,AD=AD,DD=DD). 
runGenomicPredictions<-function(blups,modelType,grms,ncores=1,gid="GID",...){
  require(sommer); 
  runOnePred<-possibly(function(trainingdata,modelType,grms){
    trainingdata[[paste0(gid,"a")]]<-factor(trainingdata[[gid]],levels=rownames(grms[["A"]]))
    if(modelType %in% c("AD","ADE")){ trainingdata[[paste0(gid,"d")]]<-factor(trainingdata[[gid]],levels=rownames(grms[["D"]])) 
    if(modelType=="ADE"){
      #trainingdata[[paste0(gid,"aa")]]<-factor(trainingdata[[gid]],levels=rownames(grms[["AA"]]))
      trainingdata[[paste0(gid,"ad")]]<-factor(trainingdata[[gid]],levels=rownames(grms[["AD"]]))
      diag(grms[["AD"]])<-diag(grms[["AD"]])+1e-06
      #trainingdata[[paste0(gid,"dd")]]<-factor(trainingdata[[gid]],levels=rownames(grms[["DD"]]))
    }
    }
    # Set-up random model statements
    randFormula<-paste0("~vs(",gid,"a,Gu=A)")
    if(modelType %in% c("AD","ADE")){
      randFormula<-paste0(randFormula,"+vs(",gid,"d,Gu=D)")
      if(modelType=="ADE"){
        randFormula<-paste0(randFormula,"+vs(",gid,"ad,Gu=AD)")
                            # "+vs(",gid,"aa,Gu=AA)",
                            # "+vs(",gid,"dd,Gu=DD)")
      }
    }
    # Fit genomic prediction model  
    fit <- mmer(fixed = drgBLUP ~1,
                random = as.formula(randFormula),
                weights = WT,
                data=trainingdata)
    # Gather the BLUPs
    gblups<-tibble(GID=as.character(names(fit$U[[paste0("u:",gid,"a")]]$drgBLUP)),
                   GEBV=as.numeric(fit$U[[paste0("u:",gid,"a")]]$drgBLUP))
    if(modelType %in% c("AD","ADE")){
      gblups %<>% mutate(GEDD=as.numeric(fit$U[[paste0("u:",gid,"d")]]$drgBLUP))
      if(modelType=="ADE"){
        gblups %<>% mutate(#GEEDaa=as.numeric(fit$U[[paste0("u:",gid,"aa")]]$drgBLUP),
                           GEEDad=as.numeric(fit$U[[paste0("u:",gid,"ad")]]$drgBLUP))
                           #GEEDdd=as.numeric(fit$U[[paste0("u:",gid,"dd")]]$drgBLUP)) 
      }
    }
    # Calc GETGVs 
    ## Note that for modelType=="A", GEBV==GETGV
    gblups %<>% 
      mutate(GETGV=rowSums(.[,grepl("GE",colnames(.))])) 
    varcomps<-summary(fit)$varcomp
    out<-tibble(gblups=list(gblups),varcomps=list(varcomps))
    return(out) 
  },otherwise = NA)
  ## Run models across all train-test splits
  ## Parallelize 
  require(furrr); plan(multiprocess); options(mc.cores=ncores);
  predictions<-blups %>% 
      mutate(genomicPredOut=future_map(TrainingData,~runOnePred(trainingdata=.,
                                                                modelType=modelType,grms=grms)))
  return(predictions) 
}
```

cbsulm18 (88 cores; 512GB) 

Model A
```{r, eval=F}
options(future.globals.maxSize= 1500*1024^2)
predModelA<-runGenomicPredictions(blups,modelType="A",grms=list(A=A),gid="GID",ncores=1)
saveRDS(predModelA,file = here::here("output","genomicPredictions_ModelA_NRCRI_2020April27.rds"))
```

Model ADE
```{r, eval=F}
options(future.globals.maxSize= 3000*1024^2)
predModelADE<-runGenomicPredictions(blups,modelType="ADE",grms=list(A=A,D=D,AD=AD),gid="GID",ncores=20)
saveRDS(predModelADE,file = here::here("output","genomicPredictions_ModelADE_NRCRI_2020April27.rds"))
```

## Plot Predictions
```{r}
library(tidyverse); library(magrittr);
predModelA<-readRDS(file = here::here("output","genomicPredictions_ModelA_NRCRI_2020April27.rds"))
predModelADE<-readRDS(file = here::here("output","genomicPredictions_ModelADE_NRCRI_2020April27.rds"))
```

```{r}
predModelA %>% 
  dplyr::select(Trait,Dataset,genomicPredOut) %>% 
  unnest(genomicPredOut) %>% 
  select(-varcomps) %>% 
  unnest(gblups) %>% 
  select(-GETGV) %>% rename(GEBV_A=GEBV) %>% 
  left_join(predModelADE %>% 
  mutate(islgl=map_lgl(genomicPredOut,is.logical)) %>% 
  filter(islgl==FALSE) %>% 
  dplyr::select(Trait,Dataset,genomicPredOut) %>% 
  unnest(genomicPredOut) %>% 
  select(-varcomps) %>% 
  unnest(gblups) %>% 
  select(Trait,Dataset,GID,GEBV) %>% rename(GEBV_ADE=GEBV)) %>% 
  mutate(GeneticGroup=NA,
         GeneticGroup=ifelse(grepl("NR17F",GID,ignore.case = T),"C2a",
                             ifelse(grepl("C2bF",GID,ignore.case = T),"C2b",
                                    ifelse(grepl("^F",GID) & !grepl("C1b",GID),"C1a",
                                           ifelse(grepl("C1b",GID,
                                                        ignore.case = T),"C1b","TrainingPop"))))) %>% 
  ggplot(.,aes(x=GEBV_A,y=GEBV_ADE,color=GeneticGroup)) + 
  geom_point() + theme_bw() + geom_abline(slope=1, color='darkred') + 
  facet_wrap(~Trait, scales = 'free') + 
  labs(title="Compare GEBV from A-only model to GEBV from A+D+E model")
```
```{r}
predModelADE %>% 
  mutate(islgl=map_lgl(genomicPredOut,is.logical)) %>% 
  filter(islgl==FALSE) %>% 
  dplyr::select(Trait,Dataset,genomicPredOut) %>% 
  unnest(genomicPredOut) %>% 
  select(-varcomps) %>% 
  unnest(gblups) %>% 
  select(Trait,Dataset,GID,GEBV,GETGV) %>% 
  mutate(GeneticGroup=NA,
         GeneticGroup=ifelse(grepl("NR17F",GID,ignore.case = T),"C2a",
                             ifelse(grepl("C2bF",GID,ignore.case = T),"C2b",
                                    ifelse(grepl("^F",GID) & !grepl("C1b",GID),"C1a",
                                           ifelse(grepl("C1b",GID,
                                                        ignore.case = T),"C1b","TrainingPop"))))) %>% 
  ggplot(.,aes(x=GEBV,y=GETGV,color=GeneticGroup)) + 
  geom_point() + theme_bw() + geom_abline(slope=1, color='darkred') + 
  facet_wrap(~Trait, scales = 'free') +
  labs(title="Model: ADE - Compare GEBV to GETGV")
```

## Write GEBVs
```{r predModelA}
predModelA %>% 
  dplyr::select(Trait,Dataset,genomicPredOut) %>% 
  unnest(genomicPredOut) %>% 
  select(-varcomps) %>% 
  unnest(gblups) %>% 
  select(-GETGV) %>% 
  spread(Trait,GEBV) %>% 
  mutate(GeneticGroup=NA,
         GeneticGroup=ifelse(grepl("NR17F",GID,ignore.case = T),"C2a",
                             ifelse(grepl("C2bF",GID,ignore.case = T),"C2b",
                                    ifelse(grepl("^F",GID) & !grepl("C1b",GID),"C1a",
                                           ifelse(grepl("C1b",GID,
                                                        ignore.case = T),"C1b","TrainingPop"))))) %>% 
  nest(GEBVs=-Dataset) %>% 
  mutate(write=map2(Dataset,GEBVs,~write.csv(x = .y, file = here::here("output",paste0("GEBV_NRCRI_",.x,"_ModelA_2020April27.csv")))))
  
```
```{r predModelADE}
## Format and write GEBV
predModelADE %>% 
  mutate(islgl=map_lgl(genomicPredOut,is.logical)) %>% 
  filter(islgl==FALSE) %>% 
  dplyr::select(Trait,Dataset,genomicPredOut) %>% 
  unnest(genomicPredOut) %>% 
  select(-varcomps) %>% 
  unnest(gblups) %>% 
  select(Trait,Dataset,GID,GEBV) %>% 
  spread(Trait,GEBV) %>% 
  mutate(GeneticGroup=NA,
         GeneticGroup=ifelse(grepl("NR17F",GID,ignore.case = T),"C2a",
                             ifelse(grepl("C2bF",GID,ignore.case = T),"C2b",
                                    ifelse(grepl("^F",GID) & !grepl("C1b",GID),"C1a",
                                           ifelse(grepl("C1b",GID,
                                                        ignore.case = T),"C1b","TrainingPop"))))) %>% 
  nest(GEBVs=-Dataset) %>% 
  mutate(write=map2(Dataset,GEBVs,~write.csv(x = .y, file = here::here("output",paste0("GEBV_NRCRI_",.x,"_ModelADE_2020April27.csv")))))
## Format and write GETGV
predModelADE %>% 
  mutate(islgl=map_lgl(genomicPredOut,is.logical)) %>% 
  filter(islgl==FALSE) %>% 
  dplyr::select(Trait,Dataset,genomicPredOut) %>% 
  unnest(genomicPredOut) %>% 
  select(-varcomps) %>% 
  unnest(gblups) %>% 
  select(Trait,Dataset,GID,GETGV) %>% 
  spread(Trait,GETGV) %>% 
  mutate(GeneticGroup=NA,
         GeneticGroup=ifelse(grepl("NR17F",GID,ignore.case = T),"C2a",
                             ifelse(grepl("C2bF",GID,ignore.case = T),"C2b",
                                    ifelse(grepl("^F",GID) & !grepl("C1b",GID),"C1a",
                                           ifelse(grepl("C1b",GID,
                                                        ignore.case = T),"C1b","TrainingPop"))))) %>% 
  nest(GETGVs=-Dataset) %>% 
  mutate(write=map2(Dataset,GETGVs,
                    ~write.csv(x = .y, 
                               file = here::here("output",paste0("GETGV_NRCRI_",.x,"_ModelADE_2020April27.csv")))))

```


# Next step

6. [Estimate genetic gain](06-GetGainEst.html)

```{bash, eval=F, include=F}
rsync --update --archive --verbose /workdir/marnin/NRCRI_2020GS/ mw489@cbsulm13.biohpc.cornell.edu:/workdir/mw489/NRCRI_2020GS;
rsync --update --archive --verbose /workdir/marnin/NRCRI_2020GS/ mw489@cbsulm15.biohpc.cornell.edu:/workdir/mw489/NRCRI_2020GS;
rsync --update --archive --verbose /workdir/marnin/NRCRI_2020GS/ mw489@cbsulm18.biohpc.cornell.edu:/workdir/mw489/NRCRI_2020GS;

rsync --update --archive --verbose /workdir/mw489/NRCRI_2020GS/ mw489@cbsurobbins.biohpc.cornell.edu:/workdir/marnin/NRCRI_2020GS;
```
