---
title: "Check prediction accuracy"
site: workflowr::wflow_site
date: "2020-October-15"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = F, tidy = T)
```

# Previous step

2.  [Get BLUPs combining all trial data](07-GetBLUPs.html): Combine data from all trait-trials to get BLUPs for downstream genomic prediction.

    -   Fit mixed-model to multi-trial dataset and extract BLUPs, de-regressed BLUPs and weights. Include two rounds of outlier removal.

# Objective

**Current Step:**

3.  [Check prediction accuracy](08-CrossValidation.html): Evaluate prediction accuracy with cross-validation.

    -   Compare prediction accuracy with vs. without IITA's training data to augment.

5-fold cross-validation. Replicate 5-times.

2 genomic models:

1.  Additive-only (**A**)
2.  Addtitive plus dominance plus additive-by-dominance epistasis (**ADE**)

# Prep. genomic data

## Get SNP data from FTP

The data for the next step can be found on the cassavabase FTP server [here](ftp://ftp.cassavabase.org/marnin_datasets/NGC_BigData/).

Can be loaded directly to R from FTP.

**NOTICE:** You need enough RAM and a stable network connection. I do the next steps, including cross-validation on a server with plenty of RAM and a good, stable network connection, rather than on my personal computer (a laptop with 16 GB RAM).

The outputs (kinship matrices and filtered snp dosages) of the steps below, which are too large for GitHub, can be found on the cassavabase FTP server [here](ftp://ftp.cassavabase.org/marnin_datasets/NRCRI_2020GS/).

```{bash, eval=F}
# activate multithread OpenBLAS for fast compute of SigmaM (genotypic var-covar matrix)
export OMP_NUM_THREADS=56
```

```{r, eval=F}
library(tidyverse); library(magrittr); 
snps<-readRDS(file=url(paste0("ftp://ftp.cassavabase.org/marnin_datasets/NGC_BigData/",
                              "DosageMatrix_RefPanelAndGSprogeny_ReadyForGP_73019.rds")))
snps5510<-readRDS(here::here("output","DosageMatrix_DCas20_5510_WA_REFimputedAndFiltered.rds"))
snps5440<-readRDS(here::here("output","DosageMatrix_DCas20_5440_WA_REFimputedAndFiltered.rds"))
rownames(snps5510) # NR19C3aF1P0002....
rownames(snps5440) # NR18S1P0082...

snps2keep<-colnames(snps) %>% 
  .[. %in% colnames(snps5510)] %>% 
  .[. %in% colnames(snps5440)]

snps<-rbind(snps[,snps2keep],
            snps5510[,snps2keep]) %>% 
  rbind(.,snps5440[,snps2keep])
gc()
#rm(list=(ls() %>% grep("snps",.,value = T, invert = T)))
```

```{r, eval=F}
blups_nrcri<-readRDS(file=here::here("output","nrcri_blupsForModelTraining_twostage_asreml_2020Oct13.rds"))
# copy or download IITA BLUPs from GitHub to data/ directory 
# https://wolfemd.github.io/IITA_2020GS/output/iita_blupsForModelTraining_twostage_asreml.rds
blups_iita<-readRDS(file=here::here("data","iita_blupsForModelTraining_twostage_asreml.rds"))
blups_nrcri %<>% 
  select(Trait,blups) %>% 
  unnest(blups) %>% 
  select(-`std error`) %>% 
  filter(GID %in% rownames(snps))
table(unique(blups_nrcri$GID) %in% rownames(snps)) # 2751
blups_iita %<>% 
  select(Trait,blups) %>% 
  unnest(blups) %>% 
  select(-`std error`) %>% 
  filter(GID %in% rownames(snps),
         !grepl("TMS13F|TMS14F|TMS15F|2013_|TMS16F|TMS17F|TMS18F",GID))
table(unique(blups_iita$GID) %in% rownames(snps)) # 1234
```

```{r, eval=F}
samples2Keep<-union(blups_nrcri$GID,blups_iita$GID) %>% 
  union(.,grep("c2|c3",rownames(snps),value = T, ignore.case = T)) %>% 
  union(.,rownames(snps5510)) %>% 
  union(.,rownames(snps5440))
length(samples2Keep) # 7423
snps<-snps[samples2Keep,]; 
gc()
```

## MAF\>1% filter

```{r, eval=F}
source(here::here("code","gsFunctions.R"))
snps %<>% maf_filter(.,0.01)
dim(snps) # [1] 7423 30220
```

## Make Add, Dom and Epi kinships

Going to use my own kinship function.

Make the kinships.

Below e.g. `A*A` makes a matrix that approximates additive-by-additive epistasis relationships.

```{r, eval=F}
A<-kinship(snps,type="add")
D<-kinship(snps,type="dom")
AD<-A*D

saveRDS(snps,file=here::here("output","DosageMatrix_NRCRI_2020Oct15.rds"))
saveRDS(A,file=here::here("output","Kinship_A_NRCRI_2020Oct15.rds"))
saveRDS(D,file=here::here("output","Kinship_D_NRCRI_2020Oct15.rds"))
saveRDS(AD,file=here::here("output","Kinship_AD_NRCRI_2020Oct15.rds"))
#rm(snps); gc()
```

**NOTICE:** The outputs (kinship matrices and filtered snp dosages) of the steps below, which are too large for GitHub, can be found on the cassavabase FTP server [here](ftp://ftp.cassavabase.org/marnin_datasets/NRCRI_2020GS/).

# Cross-validation

```{bash, eval=F}
cd /home/jj332_cas/marnin/NRCRI_2020GS/; 
export OMP_NUM_THREADS=56 # activate multithread OpenBLAS 
```

## Set-up training-testing data

```{r}
rm(list=ls())
library(tidyverse); library(magrittr); 
source(here::here("code","gsFunctions.R"))
blups_nrcri<-readRDS(file=here::here("output","nrcri_blupsForModelTraining_twostage_asreml_2020Oct13.rds"))
blups_iita<-readRDS(file=here::here("data","iita_blupsForModelTraining_twostage_asreml.rds"))

A<-readRDS(file=here::here("output","Kinship_A_NRCRI_2020Oct15.rds"))
blups_nrcri %<>% 
  select(Trait,blups) %>% 
  unnest(blups) %>% 
  select(-`std error`) %>% 
  filter(GID %in% rownames(A))
blups_iita %<>% 
  select(Trait,blups) %>% 
  unnest(blups) %>% 
  select(-`std error`) %>% 
  filter(GID %in% rownames(A),
         !grepl("TMS13F|TMS14F|TMS15F|2013_|TMS16F|TMS17F|TMS18F",GID))
# Set-up a grouping variable for:
## nrTP, C1a, C1b and C2a.
## Nest by Trait.
c1a<-blups_nrcri$GID %>% 
  unique %>% 
  grep("c1a",.,value = T,ignore.case = T) %>% 
  union(.,blups_nrcri$GID %>% unique %>% 
          grep("^F",.,value = T,ignore.case = T) %>% 
          grep("c1b",.,value = T,ignore.case = T,invert = T))
c1b<-blups_nrcri$GID %>% unique %>% grep("c1b",.,value = T,ignore.case = T)
c2a<-blups_nrcri$GID %>% unique %>% 
  grep("C2a",.,value = T,ignore.case = T) %>% 
  grep("NR17",.,value = T,ignore.case = T)
c2b<-blups_nrcri$GID %>% unique %>% 
  grep("C2b",.,value = T,ignore.case = T) %>% 
  .[!. %in% c(c1a,c1b,c2a)]
nrTP<-setdiff(unique(blups_nrcri$GID),unique(c(c1a,c1b,c2a,c2b)))
cv2do<-blups_nrcri %>%
  mutate(Group=case_when(GID %in% nrTP ~ "nrTP",
                         GID %in% c1a ~ "C1a",
                         GID %in% c1b ~ "C1b",
                         GID %in% c2a ~ "C2a",
                         GID %in% c2b ~ "C2b")) %>% 
  nest(TrainTestData=-Trait) %>% 
  left_join(blups_iita %>% 
               nest(augmentTP=-Trait))
cv2do %>% rmarkdown::paged_table()
```

```{r}
cv2do$TrainTestData[[6]] %>%
  count(Group) %>% rmarkdown::paged_table()
```

```{r}
cv2do$TrainTestData[[6]] %>% head %>% rmarkdown::paged_table()
```

The function below `runCrossVal()` function implements nfold cross-validation. Specifically, for each of **nrepeats** it splits the data into **nfolds** sets according to **gid**. So if `nfolds=5` then the the clones will be divided into 5 groups and 5 predictions will be made. In each prediction, 4/5 of the clones will be used to predict the remaining 1/5. Accuracy of the model is measured as the correlation between the BLUPs (adj. mean for each CLONE) in the *test set* and the GEBV (the prediction made of each clone when it was in the test set).

Below, 4 chunks of "5 reps x 5-fold" cross-validation are run on 1 large memory Cornell CBSU machine each (e.g. [cbsulm16; 112 cores, 512 GB RAM](https://biohpc.cornell.edu/lab/hardware.aspx)).

## CV - modelType="A"

### NRCRI alone

```{r, eval=F}
starttime<-proc.time()[3]
cv_A_nrOnly<-cv2do %>% 
  mutate(CVresults=map(TrainTestData,~runCrossVal(TrainTestData=.,
                                                  modelType="A",
                                                  grms=list(A=A),
                                                  byGroup=TRUE,augmentTP=NULL,
                                                  nrepeats=5,nfolds=5,ncores=4,gid="GID")))
runtime<-proc.time()[3]-starttime; runtime

cv_A_nrOnly %<>% mutate(Dataset="NRalone",modelType="A") %>% dplyr::select(-TrainTestData,-augmentTP)
saveRDS(cv_A_nrOnly,file=here::here("output","cvresults_A_nrOnly_2020Oct15.rds"))
```

### IITA augmented

For this one, try with `ncores=1` instead of `ncores=10`.

```{r, eval=F}
starttime<-proc.time()[3]
cv_A_iitaAugmented<-cv2do %>% 
  mutate(isnullAugment=map_lgl(augmentTP,~is.null(.))) %>% 
  filter(!isnullAugment) %>% 
  select(-isnullAugment) %>% 
  mutate(CVresults=map2(TrainTestData,augmentTP,~runCrossVal(TrainTestData=.x,
                                                  modelType="A",
                                                  grms=list(A=A),
                                                  byGroup=TRUE,augmentTP=.y,
                                                  nrepeats=5,nfolds=5,ncores=1,gid="GID")))
runtime<-proc.time()[3]-starttime; runtime
cv_A_iitaAugmented %<>% mutate(Dataset="IITAaugmented",modelType="A") %>% dplyr::select(-TrainTestData,-augmentTP)
saveRDS(cv_A_iitaAugmented,file=here::here("output","cvresults_A_iitaAugmented_2020Oct15.rds"))
```

## CV - modelType="ADE"

### NRCRI alone

```{r, eval=F}
options(future.globals.maxSize= 3000*1024^2)
D<-readRDS(file=here::here("output","Kinship_D_NRCRI_2020Oct15.rds"))
AD<-readRDS(file=here::here("output","Kinship_AD_NRCRI_2020Oct15.rds"))
starttime<-proc.time()[3]
cv_ADE_nrOnly<-cv2do %>% 
  mutate(CVresults=map(TrainTestData,~runCrossVal(TrainTestData=.,
                                                  modelType="ADE",
                                                  grms=list(A=A,D=D,AD=AD),
                                                  byGroup=TRUE,augmentTP=NULL,
                                                  nrepeats=5,nfolds=5,ncores=10,gid="GID")))
cv_ADE_nrOnly %<>% mutate(Dataset="NRalone",modelType="ADE") %>% dplyr::select(-TrainTestData,-augmentTP)
saveRDS(cv_ADE_nrOnly,file=here::here("output","cvresults_ADE_nrOnly.rds"))
runtime<-proc.time()[3]-starttime; runtime

```

### IITA augmented

```{r, eval=F}
options(future.globals.maxSize= 3000*1024^2)
D<-readRDS(file=here::here("output","Kinship_D_NRCRI_2020Oct15.rds"))
AD<-readRDS(file=here::here("output","Kinship_AD_NRCRI_2020Oct15.rds"))
starttime<-proc.time()[3]
cv_ADE_iitaAugmented<-cv2do %>% 
  mutate(isnullAugment=map_lgl(augmentTP,~is.null(.))) %>% 
  filter(!isnullAugment) %>% 
  dplyr::select(-isnullAugment) %>% 
  mutate(CVresults=map2(TrainTestData,augmentTP,~runCrossVal(TrainTestData=.x,
                                                             modelType="ADE",
                                                             grms=list(A=A,D=D,AD=AD),
                                                             byGroup=TRUE,augmentTP=.y,
                                                             nrepeats=5,nfolds=5,ncores=10,gid="GID")))
cv_ADE_iitaAugmented %<>% mutate(Dataset="IITAaugmented",modelType="ADE") %>% dplyr::select(-TrainTestData,-augmentTP)
saveRDS(cv_ADE_iitaAugmented,file=here::here("output","cvresults_ADE_iitaAugmented_2020Oct15.rds"))
runtime<-proc.time()[3]-starttime; runtime

```

## CV - modelType="AD"

### IITA augmented

```{r, eval=F}
options(future.globals.maxSize= 3000*1024^2)
D<-readRDS(file=here::here("output","Kinship_D_NRCRI_2020Oct15.rds"))
#AD<-readRDS(file=here::here("output","Kinship_AD_NRCRI_2020Oct15.rds"))
starttime<-proc.time()[3]
cv_AD_iitaAugmented<-cv2do %>% 
  mutate(isnullAugment=map_lgl(augmentTP,~is.null(.))) %>% 
  filter(!isnullAugment) %>% 
  dplyr::select(-isnullAugment) %>% 
  mutate(CVresults=map2(TrainTestData,augmentTP,~runCrossVal(TrainTestData=.x,
                                                             modelType="AD",
                                                             grms=list(A=A,D=D),
                                                             byGroup=TRUE,augmentTP=.y,
                                                             nrepeats=5,nfolds=5,ncores=4,gid="GID")))
cv_AD_iitaAugmented %<>% mutate(Dataset="IITAaugmented",modelType="AD") %>% dplyr::select(-TrainTestData,-augmentTP)
saveRDS(cv_AD_iitaAugmented,file=here::here("output","cvresults_AD_iitaAugmented_2020Oct15.rds"))
runtime<-proc.time()[3]-starttime; runtime

```

# Results

See [Results](10-Results.html)

# Next step

4.  [Genomic prediction](09-GetGBLUPs.html): Predict *genomic* BLUPs (GEBV and GETGV) for all selection candidates using all available data.
